---
title: "Homework 3"
format: html
editor: visual
---

## Task 1: Conceptual Questions

### 1. If your working directory is myfolder/homework/, what relative path would you specify to get the file located at myfolder/MyData.csv?

The relative path that you would specify to get the file is "../MyData.csv".

### 2. What are the major benefits of using R projects?

The major benefits of using R projects are:

- R projects provide a straightforward way to divide your work into multiple contexts. Each with their own:

  - Working directory

  - Workspace (environment with stored data/objects)

  - History

  - Folder structure & source documents

### 3. What is Git and what is GitHub?

Git is a version control software.

Github is an online hosting service for Git-based projects.

### 4. What are the two main differences between a tibble and a data.frame?

- Data frames match the prefix of any variable names (**partial matching**) and don’t complain if a column doesn’t exist. Tibbles are more strict: they only ever match variable names exactly and they will generate a warning if the column you are trying to access doesn’t exist.

- Tibble is a modern re-imagining of the data frame, keeping what time has proven to be effective, and throwing out what it has not. Tibbles are data.frames that are lazy and surly: they do less and complain more forcing you to confront problems earlier, typically leading to cleaner, more expressive code.

### 5. Rewrite the following nested function call using baseR’s chaining operator:

```{r}

##original 
library(tidyverse)
arrange(filter(select(as_tibble(iris), starts_with("Petal"), Species), Petal.Length <
                 1.55), Species)

##rewrited
library(tidyverse)

as_tibble(iris) |>
  select(starts_with("Petal"), Species) |>
  filter(Petal.Length < 1.55) |>
  arrange(Species)
```

### 6. What is meant by long format data and wide format data? Which do we generally prefer for statistical analysis?

**Long format data** is where each row consists of one observation and each column one variable. **Wide format** data is where we may have more than one observation in a given row. For statistical analysis we prefer **long format data** as almost all of the plotting, summarizing, and modeling we do in statistics expects data to be in long form.

## Task 2: Reading Delimited Data

The first data set is called glass.data. This data set is comma delimited. The second data set is called yeast data. This data set is white-space delimited.

```{r}
library(tidyverse)

# Read data and add column names
glass_cols <- c("Id", "RI", "Na", "Mg", "Al", "Si", "K", "Ca", "Ba", "Fe", "Type_of_glass")
glass_data <- read_csv("https://www4.stat.ncsu.edu/~online/datasets/glass.data", col_names = glass_cols)

# Print the tibble
glass_data 

# Create a vector of the names in order (1 to 7)
glass_names <- c("building_windows_float_processed", 
                 "building_windows_non_float_processed", 
                 "vehicle_windows_float_processed", 
                 "vehicle_windows_non_float_processed", 
                 "containers", 
                 "tableware", 
                 "headlamps")

glass_data |>
  # Replace the number with the corresponding name from the vector
  mutate(Type_of_glass = glass_names[Type_of_glass]) |>
  # Convert to a factor 
  mutate(Type_of_glass = as.factor(Type_of_glass)) |>
  # Filter the data
  filter(Fe < 0.2, Type_of_glass %in% c("tableware", "headlamps"))

# Read data and add column names
yeast_cols <- c("seq_name", "mcg", "gvh", "alm", "mit", "erl", "pox", "vac", "nuc", "class")
yeast_data <- read_table("https://www4.stat.ncsu.edu/~online/datasets/yeast.data", col_names = yeast_cols)

# Print the tibble
yeast_data

# Chain to remove columns and add mean/median for numeric variables
yeast_data |>
  select(-seq, -nuc) |>
  group_by(class) |>
  mutate(across(c(mcg, gvh, alm, mit, erl, pox, vac), 
                list(mean = mean, median = median), 
                .names = "{.col}_{.fn}"))
```

```{r}
library(readxl)

# Import white wine data from the first sheet
white_wine <- read_excel("white-wine.xlsx", sheet = 1)
white_wine

# Read column names from the second sheet and overwrite names
wine_names <- read_excel("white-wine.xlsx", sheet = 2) |> pull(1)
colnames(white_wine) <- wine_names

# Add column indicating 'white'
white_wine <- white_wine |> mutate(type = "white")

# Read semi-colon delimited red wine data, update names, and add type
red_wine <- read_delim("https://www4.stat.ncsu.edu/~online/datasets/red-wine.csv", delim = ";")
colnames(red_wine) <- wine_names
red_wine <- red_wine |> mutate(type = "red")

# Combine the data sets
combined_wine <- bind_rows(white_wine, red_wine)

# 6. - 9. Chain to filter, sort, select, and mutate the combined data
combined_wine |>
  filter(quality > 6.5, alcohol < 132) |>
  arrange(desc(quality)) |>
  select(contains("acid"), alcohol, type, quality) |>
  group_by(quality) |>
  mutate(mean_alcohol = mean(alcohol), sd_alcohol = sd(alcohol))
```

```{r}
library(DBI)
library(RSQLite)

# Connect to database and look at tables
con <- dbConnect(SQLite(), "Lahman.db")
dbListTables(con)

# Return 2015 Teams data using tbl() and dplyr
tbl(con, "Teams") |> 
  filter(yearID == 2015)

# Return 2015 Teams data using sql() within tbl()
tbl(con, sql("SELECT * FROM Teams WHERE yearID = 2015"))

# Return Hall of Fame players, year voted, and category
hof_players <- tbl(con, "HallOfFame") |> 
  select(playerID, yearID, category)

# Combine with People table to add first and last names
hof_people <- hof_players |>
  inner_join(tbl(con, "People") |> select(playerID, nameFirst, nameLast), by = "playerID")

# Retrieve managers, summarize records, and create win/loss percentage
managers_summary <- tbl(con, "Managers") |>
  select(playerID, G, W, L) |>
  group_by(playerID) |>
  summarize(G_managed = sum(G, na.rm = TRUE),
            Total_W = sum(W, na.rm = TRUE),
            Total_L = sum(L, na.rm = TRUE)) |>
  collect() |>
  mutate(win_loss_pct = Total_W / G_managed) |>
  arrange(desc(win_loss_pct))

# Return information for Hall of Fame inductees who managed a team
hof_people |>
  collect() |>
  inner_join(managers_summary, by = "playerID")
```
