---
title: "Homework 3"
format: html
editor_options: 
  chunk_output_type: console
---

# Preamble
This assignment practices reading, combining, and manipulating different data types, such as Excel, comma-separated values (CSV), and databases.

# Task 1: Conceptual Questions
1. If your working directory is `myfolder/homework/`, what _relative_ path would you specify to get the file located at `myfolder/MyData.csv`?

   I would use the command `../MyData.csv` to access the file.

2. What are the m ajor benefits of using `R` projects?

   It maintains a consistent working directory.

3. What is `Git` and what is `GitHub`?

   `Git` is a code version manager and `GitHub` is an online `Git` center.
    
4. What are the two main differences between a `tibble` and a `data.frame`?

   A `tibble` is "lazy and surly,"
    
5. Rewrite the following nested function call using `baseR`'s chaining operator:

```{r hw_3_task_1_question_5}
#| warning: false
#| message: false

library(tidyverse)
arrange(filter(select(as_tibble(iris), starts_with("Petal"), Species),
              Petal.Length < 1.55), Species)
```

   The appropriate code using the chaining operator is
   
```{r hw_3_task_1_question_5_response}
#| warning: false
#| message: false

library(tidyverse)
as_tibble(iris) |>
  select(starts_with("Petal"), Species) |>
  filter(Petal.Length < 1.55) |>
  arrange(Species)
```

6. What is meant by long format data and wide format data? Which do we generally prefer for statistical analysis?

   Wide format data is useful when a variable has several observations, so called because it tends to have fewer rows but more columns (one per observation) than the long format, which splits each of the observations across a different row, thus resulting in more rows but fewer columns. While the wide format is used for ease of viewing, for statistical analysis, the long format is preferred.