---
title: "Homework 3"
format: html
editor_options: 
  chunk_output_type: console
---

# Preamble
This assignment practices reading, combining, and manipulating different data types, such as Excel, comma-separated values (CSV), and databases.

# Task 1: Conceptual Questions
1. If your working directory is `myfolder/homework/`, what _relative_ path would you specify to get the file located at `myfolder/MyData.csv`?

   I would use the command `../MyData.csv` to access the file.

2. What are the major benefits of using `R` projects?

   It maintains a consistent working directory.

3. What is `Git` and what is `GitHub`?

   `Git` is a version management software and `GitHub` is an online `Git` center.
    
4. What are the two main differences between a `tibble` and a `data.frame`?

   A `tibble` is "lazy and surly,"
    
5. Rewrite the following nested function call using `baseR`'s chaining operator:

```{r hw_3_task_1_question_5}
#| warning: false
#| message: false

library(tidyverse)
arrange(filter(select(as_tibble(iris), starts_with("Petal"), Species),
              Petal.Length < 1.55), Species)
```

   The appropriate code using the chaining operator is
   
```{r hw_3_task_1_question_5_response}
#| warning: false
#| message: false

library(tidyverse)
as_tibble(iris) |>
  select(starts_with("Petal"), Species) |>
  filter(Petal.Length < 1.55) |>
  arrange(Species)
```

6. What is meant by long format data and wide format data? Which do we generally prefer for statistical analysis?

   Wide format data is useful when a variable has several observations, so called because it tends to have fewer rows but more columns (one per observation) than the long format, which splits each of the observations across a different row, thus resulting in more rows but fewer columns. While the wide format is used for ease of viewing, for statistical analysis, the long format is preferred.
   
# Task 2: Reading Delimited Data
The data sets for this portion come from the [UCI machine learning repository](https://archive.ics.uci.edu/datasets).

## Glass Data
Opening this file shows that the data is delimited by commas, `,`. The source of the data is [here]( https://www4.stat.ncsu.edu/~online/datasets/glass.data).

1. First we create the column names and read in the data.
```{r hw_3_task_2_response_1}
glass_col_names <- c("Id", "RI", "Na", "Mg", "Al", "Si", "K", "Ca", "Ba", "Fe",
                     "Type_of_glass")

glass <- read_csv("https://www4.stat.ncsu.edu/~online/datasets/glass.data",
                  col_names = glass_col_names
                  )

glass
```

2. Next we use a chain to overwrite the `Type_of_glass` variable and create a factor version of this variable. Then we replace the numbered factors with descriptive string factors.
```{r hw_3_task_2_response_2}
glass_factors <- c(building_windows_float_processed = '1', 
                   building_windows_non_float_processed = '2',
                   vehicle_windows_float_processed = '3',
                   vehicle_windows_non_float_processed = '4',
                   containers = '5',
                   tableware = '6',
                   headlamps = '7')

glass <- glass |> 
            mutate("Type_of_glass" = as.factor(Type_of_glass)) |>
            mutate("Type_of_glass" = fct_recode(Type_of_glass,
                                                !!!glass_factors))

glass
```

Note that while a warning is given regarding the factor level `4`, viz., `vehicle_windows_non_float_processed`, this is merely because this factor, as indicated in the file description, has no observations in this database.

3. Finally, we purge the database of all observations except those where the iron, `Fe`, is less than $0.2$ and the `Type_of_glass` is either `tableware` or `headlamps`.
```{r hw_3_task_2_response_3}
glass <- glass |>
            filter((Fe < 0.3) & 
                      (Type_of_glass %in% c("tableware", "headlamps")))

glass
```

## Yeast Data
This data set, found [here](https://www4.stat.ncsu.edu/online/datasets/yeast.data), is delimited by ~~double-spaces, `  ` a tab, evidently~~ white spaces, but some variables have spaces within them, so we must use the `read_table` function, which separates columns across __any__ whitespace.

1. First we read in the data, manually assigned column names.
```{r hw_3_task_2_response_4}
yeast_columns <- c("seq_name", "mcg", "gvh", "alm", "mit", "erl", "pox", "vac",
                   "nuc", "class")

yeast <- read_table("https://www4.stat.ncsu.edu/online/datasets/yeast.data",
                    col_names = yeast_columns)

yeast
```

2. Next we remove the `seq_name` and `nuc` columns.
```{r hw_3_task_2_response_5}
yeast <- yeast |>
            select(-seq_name, -nuc)

yeast
```

3. Finally we add columns corresponding to the mean and median of the numeric variables, `mcg`, `gvh`, `alm`, `mit`, `erl`, and `pox` at each class grouping.
```{r hw_3_task_2_response_6}
yeast <- yeast |>
            group_by(class) |>
            mutate(across(mcg:vac,
                          ~ mean(.x, na.rm = TRUE), # this part was tricky, .x is a filler for the generic row
                          .names = "mean_{.col}")) |>
            mutate(across(mcg:vac,
                          ~ median(.x, na.rm = TRUE),
                          .names = "med_{.col}"))

yeast
```

_NB_: The above was more complicated than it needed to be, accounting for `NA` values which did not exist in the original data set. They were introduced by incorrectly reading in the file with `read_delim` and using a double-space, `  `, as the delimiter. Some of the observations had a space in the `seq_name`, causing the `mcg` column to begin with a space and thus interpreted as an `NA`.

# Task 3: Combining Excel and Delimited Data
